% T I T L E   P A G E
% -------------------
% Last updated May 24, 2011, by Stephen Carr, IST-Client Services
% The title page is counted as page `i' but we need to suppress the
% page number.  We also don't want any headers or footers.
\pagestyle{empty}
\pagenumbering{roman}

% The contents of the title page are specified in the "titlepage"
% environment.
\begin{titlepage}
  \begin{center}
  \vspace*{1.0cm}

  \Huge
  %% {\bf Methods for Shallow and Deep Learning in\\Spiking Neural Networks}
  {\bf Spiking Deep Neural Networks:\\Engineered and Biological\\Approaches to Object Recognition}
  %% {\bf Engineered and Biological\\Spiking Deep Neural Networks\\for Object Recognition}

  \vspace*{1.0cm}

  \normalsize
  by \\

  \vspace*{1.0cm}

  \Large
  Eric Hunsberger \\

  \vfill

  \normalsize
  A thesis \\
  presented to the University of Waterloo \\
  in fulfillment of the \\
  thesis requirement for the degree of \\
  Doctor of Philosophy \\
  in \\
  Systems Design Engineering \\

  \vspace*{2.0cm}

  Waterloo, Ontario, Canada, 2017 \\

  \vspace*{1.0cm}

  \copyright\ Eric Hunsberger 2017 \\
  \end{center}
\end{titlepage}

% The rest of the front pages should contain no headers and be numbered using Roman numerals starting with `ii'
\pagestyle{plain}
\setcounter{page}{2}

\cleardoublepage

% --- Committee membership page
%% \phantomsection
%% \addcontentsline{toc}{chapter}{Examining Committee Membership}

\noindent
\textbf{Examining Committee Membership}

\noindent
The following served on the Examining Committee for this thesis.
The decision of the Examining Committee is by majority vote.

\vspace*{0.5cm}

\noindent
\begin{tabular}{p{2in}l}
External Examiner         & Bruno Olshausen \\
  & Professor,\\
  & Helen Wills Neuroscience Institute and\\
  & School of Optometry,\\
  & University of California, Berkeley\\
\\
Supervisor                & Chris Eliasmith \\
  & Professor,\\
  & Department of Philosophy and\\
  & Department of Systems Design Engineering,\\
  & University of Waterloo\\
\\
Supervisor                & Jeff Orchard \\
  & Associate Professor,\\
  & Cheriton School of Computer Science,\\
  & University of Waterloo\\
\\
Internal Member           & Bryan Tripp \\
  & Associate Professor,\\
  & Department of Systems Design Engineering,\\
  & University of Waterloo\\
\\
Internal Member           & Alexander Wong\\
  & Associate Professor,\\
  & Department of Systems Design Engineering,\\
  & University of Waterloo\\
\\
Internal-External Member  & Steven Waslander\\
  & Associate Professor,\\
  & Department of Mechanical and\\
  & Mechatronics Engineering,\\
  & University of Waterloo\\
\end{tabular}

\cleardoublepage

% D E C L A R A T I O N   P A G E
% -------------------------------
%% \phantomsection
%% \addcontentsline{toc}{chapter}{Author's Declaration}

\noindent
%% THIS IS A DRAFT AND NOT A FINAL COPY OF THIS THESIS!
I hereby declare that I am the sole author of this thesis.
This is a true copy of the thesis, including any required final revisions,
as accepted by my examiners.

  \bigskip

\noindent
%% DO NOT DISTRIBUTE!
I understand that my thesis will be made electronically available to the public.

\cleardoublepage

% A B S T R A C T
% ---------------
\phantomsection  % proper hyperref linking
\addcontentsline{toc}{chapter}{Abstract}

\begin{center}\textbf{Abstract}\end{center}

\noindent
Modern machine learning models are beginning to rival human performance
on some realistic object recognition tasks,
but we still lack a full understanding
of how the human brain solves this same problem.
This thesis combines knowledge from machine learning and computational neuroscience
to create models of human object recognition
that are increasingly realistic
both in their treatment of low-level neural mechanisms
and in their reproduction of high-level human behaviour.

First, I present extensions to the Neural Engineering Framework
to make its preferred type of model---the ``fixed-encoding'' network---%
more accurate for object recognition tasks.
These extensions include better distributions---such as Gabor filters---for the encoding weights,
and better loss functions---namely weighted squared loss, softmax loss, and hinge loss---%
to solve for decoding weights.

Second, I introduce increased biological realism into
deep convolutional neural networks trained with backpropagation,
by training them to run using spiking leaky integrate-and-fire (LIF) neurons.
Convolutional neural networks have been successful in machine learning,
and I am able to convert them to spiking networks
while retaining similar levels of performance.
I present a novel method to smooth the LIF rate response function
in order to avoid the common problems associated with differentiating
spiking neurons in general and LIF neurons in particular.
I also derive a number of novel characterizations of spiking variability,
and use these to train spiking networks to be more robust to this variability.

Finally, to address the problems with implementing backpropagation in a biological system,
I train spiking deep neural networks using the more biological Feedback Alignment algorithm.
I examine this algorithm in depth, including many variations on the core algorithm,
methods to train using non-differentiable spiking neurons,
and some of the limitations of the algorithm.
Using these findings, I construct a spiking model that
learns online in a biologically realistic manner.

The models developed in this thesis
help to explain both how spiking neurons in the brain
work together to allow us to recognize complex objects,
and how the brain may learn this behaviour.
Their spiking nature allows them to be implemented
on highly efficient neuromorphic hardware,
opening the door to object recognition on energy-limited devices
such as cell phones and mobile robots.

\cleardoublepage
%\newpage

% A C K N O W L E D G E M E N T S
% -------------------------------
%% \phantomsection  % proper hyperref linking
%% \addcontentsline{toc}{chapter}{Acknowledgements}

\begin{center}\textbf{Acknowledgements}\end{center}

\noindent
First and foremost,
I would like to thank my supervisors,
Chris Eliasmith and Jeff Orchard.
Their guidance, ideas, and keen questions
have pushed me to look more deeply
at some of the more pressing and interesting problems in the field.

Chris has gone beyond his duties by being not only a great mentor,
but also a friend.
He has filled the CNRG lab at UW with amazing people,
and I would like to thank everyone for her or his unique contribution
to making the lab both a fertile ground for new and exciting ideas,
and a fun and awesome place to be.
I owe a special gratitude to those who first welcomed me to the lab,
and who greatly influenced my early growth as a researcher and software developer:
Trevor Bekolay, Terry Stewart, Dan Rasmussen, Travis DeWolf, Xuan Choo,
and James Bergstra.
More recently, thanks to Peter Duggins, Mariah Martin Shein, and Peter Blouw
for the numerous musical interludes
that helped me survive the last months of writing.

Many of these past and present CNRG members
also form the core development team for Nengo.
Without this outstanding piece of software,
I could not have done all the things I did in this thesis,
so a special thanks to the whole Nengo team.

I was also fortunate to have such a great committee
to help me through the whole process.
In particular, I would like to thank Brian Tripp,
who has for years provided insightful comments, feedback, and ideas
across many areas of my research.

Finally, I could not have come this far
without the continued support of my family and friends.
Thank you for being genuinely interested in what I do,
even when I am sure it sounds esoteric or downright dull.
A special thanks to Sean Anderson, Matt McGill, Matt Wilson, and Nick Fischer,
for over 75 combined years of being great friends to me.
A very special thanks my family: Brian, Jocelyn, Alex, and Marg.
And the specialest thanks to Ruth,
who has not borne the burden with me as long as some of the others,
%% but who has borne so much more of it.
but who has helped to carry so much more of it.


\cleardoublepage
%\newpage

% D E D I C A T I O N
% -------------------
%% \phantomsection  % proper hyperref linking
%% \addcontentsline{toc}{chapter}{Dedication}

\begin{center}\textbf{Dedication}\end{center}

\noindent
Half-way through this thesis project---%
which one may know is often the point when motivation begins to wane---%
I received the following fortune inside a cookie:
\begin{displayquote}
If the brain were so simple we could understand it,\\
we would be so simple we couldn't.
\end{displayquote}
Since then, I have kept this fortune on my desk to motivate me.
%% I dedicate this thesis to proving it wrong.
I dedicate this thesis to all of us
who work to prove this fortune cookie wrong.

%% Since then, I have kept it on my desk to motivate me.
%% I dedicate this thesis to proving that fortune cookie wrong.
%% May we, as researchers of the brain,
%% continue to work together to prove this fortune cookie wrong.

\cleardoublepage
%\newpage

% T A B L E   O F   C O N T E N T S
% ---------------------------------
\phantomsection  % proper hyperref linking
\addcontentsline{toc}{chapter}{Table of Contents}
\renewcommand\contentsname{Table of Contents}
\tableofcontents
\cleardoublepage
\phantomsection
%\newpage

% L I S T   O F   T A B L E S
% ---------------------------
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page
%\newpage

% L I S T   O F   F I G U R E S
% -----------------------------
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page
%\newpage

% L I S T   O F   S Y M B O L S
% -----------------------------
% To include a Nomenclature section
%% \addcontentsline{toc}{chapter}{\textbf{Nomenclature}}
%% \renewcommand{\nomname}{Nomenclature}
%% \printglossary
%% \cleardoublepage
%% \phantomsection % allows hyperref to link to the correct page
%% % \newpage

% Change page numbering back to Arabic numerals
\pagenumbering{arabic}

%%  LocalWords:  Olshausen Cheriton Tripp Waslander CNRG UW Bekolay
%%  LocalWords:  DeWolf Xuan Choo Bergstra Duggins Mariah Shein Blouw
%%  LocalWords:  McGill Marg specialest
